Component response rate variation drives stability in large complex systems

### 2018-06-01 notebook/SI.Rmd
First complete draft

```{r, echo = FALSE}
plot_evo_out <- function(evo_out){
    gammas <- get_top_evo_out(evo_out, size = 9);
    par(mfrow = c(3, 3), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
    hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[7]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xlim = c(0, 1.1));
    hist(gammas[[8]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[9]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         yaxt = "n", xlim = c(0, 1.1));
    mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
          text = expression(paste("Component response rate (",gamma,")")));
    mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
}
get_top_evo_out <- function(evo_out, size){
    highest <- max(evo_out);
    gammas  <- NULL;
    while(size > 0){
        pos <- which(evo_out == highest);
        len <- length(pos);
        nli <- NULL;
        for(i in 1:len){
            start          <- pos[i] + 1;
            end            <- pos[i] + highest;
            gammas[[size]] <- evo_out[start:end];
            size           <- size - 1;
            if(size == 0){
                break;
            }
        }
        highest <- highest - 1;
    }
    return(gammas);
}
evo_res <- read.csv(file = "sim_results/evolved/evo_results.csv");
kable(evo_res);
```

@@

The above is not nearly as nasty as it looks, especially because it is entirely reasonable to simply use convenient values to parameterise it. The variable $x^{*}$ is just the vector of species abundances at equilibrium (we need all of them to be positive). The matrix $I$ is just the identity matrix (1s on the diagonal, 0s on the off-diagonal elements), and the value $\theta$ is just the strength of intraspecific competition -- we can just set this to $\theta = -1$ (as others have). The variable $C$ is just the connectance of the community, which we will also set to 1 for convenience (lower values would mean that some species don't interact with one another, corresponding to an off-diagonal matrix element of 0). And $\delta$ just affects the strength of interactions -- which we can set to $\delta = 0$ for strong interactions. Hence, the whole $(CS)^{-\delta} = 1$, so we're just adding the diagonal matrix of -1s ($\theta I$) to $A$, which has a diagonal of all zeros and an off-diagonal effecting species interactions,

@@

To check the feasibility criteria, all that needs to be done is to invert $(\theta I + A)$ and multiply the matrix by the vector of growth rates $r$, which we can also just set to 1. So really, we're just multiplying $-(\theta I + A)$ by a vector of ones and checking to make sure that all the values are positive. This isn't much extra work, but it will probably go a long way toward satisfying any reviewers familiar with @Dougoud2018.

>>>>

```{r, echo = FALSE}
get_top_evo_out <- function(evo_out, size){
    highest <- max(evo_out);
    gammas  <- NULL;
    while(size > 0){
        pos <- which(evo_out == highest);
        len <- length(pos);
        nli <- NULL;
        for(i in 1:len){
            start          <- pos[i] + 1;
            end            <- pos[i] + highest;
            gammas[[size]] <- evo_out[start:end];
            size           <- size - 1;
            if(size == 0){
                break;
            }
        }
        highest <- highest - 1;
    }
    return(gammas);
}
plot_evo_out <- function(evo_out){
    gammas <- get_top_evo_out(evo_out, size = 9);
    par(mfrow = c(3, 3), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
    hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[7]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xlim = c(0, 1.1));
    hist(gammas[[8]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         yaxt = "n", xlim = c(0, 1.1));
    hist(gammas[[9]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         yaxt = "n", xlim = c(0, 1.1));
    mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
          text = expression(paste("Component response rate (",gamma,")")));
    mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
}
evo_res <- read.csv(file = "sim_results/evolved/evo_results.csv");
kable(evo_res);
```

@@

In the above, $x^{*}$ is the vector of species abundances at equilibrium (for feasibility, all values in $x^{*}$ must be positive). The matrix $I$ is the identity matrix (1s on the diagonal, 0s on the off-diagonal elements), and the value $\theta$ is strength of intraspecific competition (diagonal elements). As I have done elsewhere, diagonal values are set to $-1$, so $\theta = -1$. The variable $C$ is just the interconnectedness of the community, which was set to $C = 1$ throughout the manuscript and supporing information, except [where otherwise noted](#connectance). The variable $\delta$ just affects the strength of interactions, which are held in $A$. In the main text, $\delta = 0$ underlying strong interactions. Hence, the whole $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta I$) is added to $A$, which has a diagonal of all zeros and an off-diagonal effecting species interactions. The above criteria is therefore reduced to the below; note that the parenthetical in both equations produces an $M$ matrix as used throughout the main text and supplemental information.

@@

To check the feasibility criteria, I therefore inverted $M = (\theta I + A)$ and multiplied elements by -1, then multiplied the resulting matrix by the vector of population growth rates $r$. Feasibility is satisfied if all of the elements of the resulting vector are positive.

The population growth rate for an individual species $i$ is sampled from a normal distribution of $r_{i} \sim \mathcal{N}(0, 0.4^{2})$, as shown in the `rand_gen_var` function in [the section](#IncrS) on "Stability across increasing $S$" above. Hence, each component $i$ of the complex system $M$ is assumed to be a species with a growth rate of $r_{i}$. Note that negative intrinsic growth rates are not unrealistic, and will occur in obligate mutualists in the absence of a partner.

When feasibility was evaluated with and without variation in $\gamma$, there was no increase in stability for $M$ where $\gamma$ varied as compared to where $\gamma = 1$. Results below illustrate this result, which was general to all other simulations performed.

```{r, echo = FALSE}
dat      <- read.csv(file = "sim_results/C_1/random_all.csv");
dat      <- dat[,-1]; # Extra row-indicating column removed
feas_tab <- cbind(dat[,1], dat[,8:13]);
colnames(feas_tab)[1] <- "S";
kable(feas_tab);
```

Hence, in general, $Var(\gamma)$ does not appear to affect feasibility in pure species interaction networks.
